---
title: "Final Code Airline Etiquette and Behavior Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library("readxl")  
```


#Read Data  
```{r}
data= read_excel("C:/Users/grfri/Google Drive/Spring 2022/Multivariate/johnson_wichern_data/flying-etiquette.xlsx")  



#Remove observations from subjects who have never flown  

data = data[!data$`How often do you travel by plane?` =="Never",]  



#Pull out baby_rude data 

baby = data[,c('Gender', 
               'Do you have any children under 18?', 
               'In general, is itrude to bring a baby on a plane?',
               'Age',
               'Household Income',
               'How often do you travel by plane?')]   



#Remove rows with NA from dataset   

baby_cleaned = baby[complete.cases(baby),]  



#Set variables   

library(tidyverse) 

baby_cleaned = mutate(baby_cleaned,   
                      
                      children18 = ifelse(baby_cleaned$`Do you have any children under 18?` 
                                          == 'Yes', 1, 0)   ,   
                      
                      babyrude = ifelse(baby_cleaned$`In general, is itrude to bring a baby on a plane?`
== 'No, not at all rude', 0, 1),  
                      
                      Gender = ifelse(baby_cleaned$Gender 
== 'Female', 1, 0),  
                      
                      Income = factor(baby_cleaned$`Household Income`,
                                    levels = c("$0 - $24,999",
                                               "$25,000 - $49,999",
                                               "$50,000 - $99,999",
                                               "$100,000 - $149,999",
                                               "150000")),  
                      
                      Frequency= factor(baby_cleaned$'How often do you travel by plane?', 
                                        levels = c("Never", 
                                                   "Once a year or less", 
                                                   "Once a month or less", 
                                                   "A few times per month", 
                                                   "A few times per week", 
                                                   "Every day")))  



baby_cleaned$Income = as.numeric(baby_cleaned$Income)  

baby_cleaned$Frequency = as.numeric(baby_cleaned$Frequency)  



#Change into ordinal   

age = as.factor(baby_cleaned$Age)   

levels(age) <- list(A="18-29", B="30-44", C="45-60", D= "> 60")  

age = as.numeric(age)  

baby_cleaned$Age = age   
```


#Correlation matrix 
```{r}
cormatrix = cbind(baby_cleaned$Age, baby_cleaned$Income,  baby_cleaned$Frequency)  

colnames(cormatrix) =list("Age", "Income", "Frequency of flying")  

cor(cormatrix)  
```


Logistic Regression 1: Is It Rude To Bring A Baby On A Flight?
###Backward Stepwise Regression to see if any of the parameters are significant.  

#Pull out data  
```{r}
rude_babies = 
  data[,c('In general, is itrude to bring a baby on a plane?',
                      'Age', 
                      'Do you have any children under 18?', 
                      'Gender'  
                      , 'Is itrude to move to an unsold seat on a plane?',
                      'Is it rude to wake a passenger up if you are trying to go to the bathroom?',  
                      'In general, is it rude to knowingly bring unruly children on a plane?', 
                      'Is it rude to ask someone to switch seats with you in order to be closer to friends?',  
                      'How often do you travel by plane?', 
                      'Is itrude to ask someone to switch seats with you in order to be closer to family?',   
                      'Is itrude to wake a passenger up if you are trying to walk around?', 
                      'Is itrude to recline your seat on a plane?')]  



## Remove NA  

rude_babies = rude_babies[complete.cases(rude_babies),]  



##Set variables  

#Change into ordinal  

rude_babies$Age = as.factor(rude_babies$Age)    

levels(rude_babies$Age) <- list(A="18-29", B="30-44", C="45-60", D= "> 60")   

rude_babies$Age = as.numeric(rude_babies$Age)   



#Change into binary  

rude_babies$Gender = ifelse(rude_babies$Gender == 'Female', 1, 0)  

rude_babies$`Do you have any children under 18?` = 
  ifelse(rude_babies$`Do you have any children under 18?` == 'Yes', 1, 0)  

rude_babies$`In general, is itrude to bring a baby on a plane?` =
  ifelse(rude_babies$`In general, is itrude to bring a baby on a plane?` ==
           'No, not at all rude', 0, 1)  

rude_babies$`Is itrude to move to an unsold seat on a plane?` =
  ifelse(rude_babies$`Is itrude to move to an unsold seat on a plane?` ==
           "No, not rude at all", 0, 1)  

rude_babies$`Is it rude to wake a passenger up if you are trying to go to the bathroom?` = 
  ifelse(rude_babies$`Is it rude to wake a passenger up if you are trying to go to the bathroom?` == 
           "No, not at all rude", 0, 1)  

rude_babies$`In general, is it rude to knowingly bring unruly children on a plane?` = 
  ifelse(rude_babies$`In general, is it rude to knowingly bring unruly children on a plane?` ==
           "No, not at all rude", 0, 1)  

rude_babies$`Is it rude to ask someone to switch seats with you in order to be closer to friends?` = 
  ifelse(rude_babies$`Is it rude to ask someone to switch seats with you in order to be closer to friends?` ==
           "No, not at all rude", 0, 1)  

rude_babies$`How often do you travel by plane?` =
  ifelse(rude_babies$`How often do you travel by plane?`== 
                                                           "Never" |
           rude_babies$`How often do you travel by plane?` == 
                                                           "Once a year or less", 0, 1)  

rude_babies$`Is itrude to ask someone to switch seats with you in order to be closer to family?` = 
  ifelse(rude_babies$`Is itrude to ask someone to switch seats with you in order to be closer to family?` ==
           "No, not at all rude", 0, 1)  

rude_babies$`Is itrude to wake a passenger up if you are trying to walk around?` = 
  ifelse(rude_babies$`Is itrude to wake a passenger up if you are trying to walk around?` ==
           "No, not at all rude", 0, 1)  

rude_babies$`Is itrude to recline your seat on a plane?` =
  ifelse(rude_babies$`Is itrude to recline your seat on a plane?` ==
           "No, not rude at all", 0, 1)  



##GLM  

rude_babies_fit = glm(rude_babies$`In general, is itrude to bring a baby on a plane?` ~ ., 
                      family = binomial, data = rude_babies)  



##Stepwise  

library(MASS)  

stepwise = stepAIC(rude_babies_fit, direction = "backward")  



coefficients(stepwise)  



#Age  

exp(0.2832798) 
#If younger, estimated odds of finding babies rude = 32.7% higher.  

#Children under 18 

exp(1.0099736) 
# Estimated odds of passengers who did not have young children were 
#2.7 times higher than those that did 

#Gender  

exp(0.7493112) #If male, estimated odds of finding babies rude = twice that of females.  

#Wake passenger to go to bathroom  

exp(0.3661117) 
#If you believe it is rude to wake up a passenger to go to the bathroom,
#estimated odds of finding bringing babies rude is 44% higher than if you do not.  

#Rude to knowingly bring unruly children?  

exp(3.3214522) #If you think it is rude to knkowingly bring unruly children on a plane, 
#the estimated odds of finding bringing babies rude is 27.7 times that of people who do not.  

#Switch seats to be closer to family  

exp(0.6996614) #Estimated odds of finding bringing babies rude is twice as high
#if you believe it is rude to switch seats to be closer to family than not.  

#Recline seat  

exp(0.2733699) #If you believe it is rude to recline, 
#the estimated odds of finding babies rude is 31.4% higher than if you do not.   
```


#Final Baby GLM:  
```{r}
babyrude = rude_babies$`In general, is itrude to bring a baby on a plane?`  

age = rude_babies$Age  

gender = rude_babies$Gender  

wake = 
  rude_babies$`Is it rude to wake a passenger up if you are trying to go to the bathroom?`  

children = 
  rude_babies$`In general, is it rude to knowingly bring unruly children on a plane?`  

recline_rude = rude_babies$`Is itrude to recline your seat on a plane?`  

closer_family = 
  rude_babies$`Is itrude to ask someone to switch seats with you in order to be closer to family?`  

children18 = rude_babies$`Do you have any children under 18?` 



babyrude_fit = glm(babyrude ~ age + gender + wake + children + recline_rude + closer_family 
                   + children18, 
                   family = binomial, data = rude_babies)  

summary(babyrude_fit) 

#Recline_rude is included in the stepwise regression model, but its p-value is 0.1052  
```



################ Plotting Baby GLM ################ 
```{r}
stepwise_glm = ggplot(data.frame(x = c(-20, 30)), aes(x = x)) + 
  
  stat_function(fun = function(x) {exp(-3.0048714  + 0.3661117 *x)/(1 + exp(-3.0048714  
                                                                            + 0.3661117 *x))}, color = "black", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  + 3.3214522 *x)/(1 + exp(-3.0048714  + 3.3214522  *x))}, color = "red", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  + 0.2733699 *x)/(1 + exp(-3.0048714  + 0.2733699  *x))}, color = "blue", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  + 0.6996614*x)/(1 + exp(-3.0048714  + 0.6996614 *x))}, color = "forestgreen", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  - 0.2832798 *x)/(1 + exp(-3.0048714 - 0.2832798  *x))}, color = "purple", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  - 0.7493112 *x)/(1 + exp(-3.0048714  - 0.7493112  *x))}, color = "goldenrod", lwd = 1) + 
  
  stat_function(fun = function(x) 
    {exp(-3.0048714  - 1.0099736 *x)/(1 + exp(-3.0048714  - 1.0099736  *x))}, color = "cyan", lwd = 1) + 
  
  annotate("text", x = 26, y = 0.9, label = "Wake passenger", color = "black") + 
  
  annotate("text", x = 5, y = 1.03, label = "Unruly Children", color = "red") + 
  
  annotate("text", x = 21, y = 0.75, label = "Recline seat", color = "blue") + ylab("Probability") + 
  
  annotate("text", x = 20, y = 1.03, label = "Switch seat family", color = "forestgreen") + 
  
  annotate("text", x = -19, y = 0.85, label = "Age", color = "purple") + 
  
  annotate("text", x = -12, y = 0.95, label = "Gender", color = "goldenrod") + 
  
  annotate("text", x = -12, y = 1.03, label = "Children18", color = "cyan") 

stepwise_glm 
```


################ ROC curve Baby GLM ################ 
```{r}
library(pROC)  



rocplot <- roc(babyrude ~ fitted(babyrude_fit), data= rude_babies)  

plot.roc(rocplot, legacy.axes=TRUE, colorize=T) 
# Specificity on x axis if legacy.axes=F  



thePar <- par(pty = "s")  

plot.roc(rocplot, legacy.axes = TRUE, asp = F)  

auc(rocplot)  



## tidyverse  

model <- data.frame(outcome = babyrude,   
                    
                    predicted = predict(babyrude_fit, rude_babies, type = "response"))   



library(plotROC)  

ggplot(model, aes(d = outcome, m = predicted)) +   
  
  geom_roc() +   
  
  style_roc(xlab = "1-Specificity",  
            
            ylab ="Sensitivity",  
            
            minor.breaks = c(seq(0, 0.1, by = 0.02), seq(0.9, 1, by = 0.02))) +  
  
  ggtitle(" 'Is It Rude To Bring A Baby On A Flight?' ROC Curve") +  
  
  theme(aspect.ratio=1) +  
  
  geom_abline(slope = 1, color = "grey92")  

auc(rocplot) # auc = area under ROC curve = concordance index  



##AUC: 0.7371 
```


################ Is Reclining You Seat Rude? GLM 2 ################  
```{r}
#Make table for recline  

recline_data = data[,c('Gender', 
                       'Age', 
                       'How tall are you?', 
                       'How often do you travel by plane?', 
                       'Do you ever recline your seat when you fly?',
                       'Is itrude to recline your seat on a plane?' )]  



#Remove rows with NA from dataset  

recline_data = recline_data[complete.cases(recline_data),]  



#Define variables  

rec_freq = recline_data$`How often do you travel by plane?`  

rude_rec = recline_data$`Is itrude to recline your seat on a plane?`  

doyourecline = recline_data$`Do you ever recline your seat when you fly?`  



#Change variables  

recline_data$Gender = ifelse(recline_data$Gender == 'Female', 1, 0)  



recline_data$Age = as.factor(recline_data$Age)    

levels(recline_data$Age) <- list(A="18-29", B="30-44", C="45-60", D= "> 60")   

recline_data$Age = as.numeric(recline_data$Age)   



rec_freq = ifelse(rec_freq == "Never" | rec_freq == "Once a year or less", 0, 1)  

recline_data$`How often do you travel by plane?` = rec_freq  



doyourecline = ifelse(doyourecline == "Never", 0, 1)  

recline_data$`Do you ever recline your seat when you fly?` = doyourecline  



rude_rec = ifelse(rude_rec == "No, not rude at all", 0, 1)  

recline_data$`Is itrude to recline your seat on a plane?` = rude_rec  



#Separate tall into feet and inches  

library(tidyr)  

recline_data = separate(data = recline_data, 
                        col = 'How tall are you?', 
                        into = c("feet", "inches"), sep = "\\'")  



#Remove Under 5' and Above 6' since no longer continuous  

recline_data$height = ifelse(recline_data$feet == "6", 1, 0) 
#If feet = 6 = 1, if not (if 5) = 0  



#GLM 

fit = glm(`Is itrude to recline your seat on a plane?`~ 
            Age + 
            Gender + 
            height + 
            rec_freq + 
            doyourecline, 
          family = binomial, data = recline_data)   

summary(fit)   

stepwise1 = stepAIC(fit, direction = "backward")

#Stepwise keeps Age, Rec_freq, and doyourecline

fit1 = fit = glm(`Is itrude to recline your seat on a plane?`~ 
                   Age + 
                   rec_freq + 
                   doyourecline, family = binomial, data = recline_data)

summary(fit1)
#Coefficients barely changed, Age is still insignificant

exp(0.47841) #Estimated odds of subjects who fly more frequently (more than once a year) 
#are 61.4% greater than those who do not.  

exp(2.17059) #Those who do not recline their seats 
#are 8.76 times more likely to find reclining seats rude. 


model1 <- data.frame(outcome = recline_data$`Is itrude to recline your seat on a plane?`,   
                    
                    predicted = predict(fit1,recline_data, type = "response"))   



ggplot(model1, aes(d = outcome, m = predicted)) +   
  
  geom_roc() +   
  
  style_roc(xlab = "1-Specificity",  
            
            ylab ="Sensitivity",  
            
            minor.breaks = c(seq(0, 0.1, by = 0.02), seq(0.9, 1, by = 0.02))) +  
  
  ggtitle(" 'Is It Rude To Recline Your Seat On A Plane?' ROC Curve") +  
  
  theme(aspect.ratio=1) +  
  
  geom_abline(slope = 1, color = "grey92")  
rocplot1 <- roc(`Is itrude to recline your seat on a plane?` ~
                  fitted(fit1), data= recline_data) 
auc(rocplot1) # auc = 0.6995

```


Contingecy Tables:


################ Contingency table between rudeness of reclining and rudeness of  baby ################ 
```{r}
library(PropCIs) 





RecB1 = data[,c('Is itrude to recline your seat on a plane?',
                'In general, is itrude to bring a baby on a plane?')]   

RecB1 = RecB1[complete.cases(RecB1),]  



RecB1$`Is itrude to recline your seat on a plane?` = 
  ifelse(RecB1$`Is itrude to recline your seat on a plane?` ==
           "No, not rude at all", "No", "Yes") 

RecB1$`In general, is itrude to bring a baby on a plane?` = 
  ifelse(RecB1$`In general, is itrude to bring a baby on a plane?`  == 
           'No, not at all rude', "No", "Yes") 



RecB2 = cbind(RecB1$`Is itrude to recline your seat on a plane?`, 
              RecB1$`In general, is itrude to bring a baby on a plane?`) 

colnames(RecB2) <- c("Reclining_Rude", "Baby_Rude")  

cont= xtabs(~  Baby_Rude + Reclining_Rude,   RecB2)  

cont #Contingency Table  

mcnemar.test(cont, correct=F) #test statistic 24.925, df = 1, p-value = 5.96e-07 



diffpropci.Wald.mp(127,220,849, 0.95)  



###(n12, n21, n , ci level) #Wald confidence difference b/w marginals  

####Probability of finding a baby on a plane rude is 6-15% higher 
###than finding reclining rude  
```


#################### Tall vs Recline #################### 
```{r}
library(tidyverse) 



recline = data[,c('How tall are you?', 'Is itrude to recline your seat on a plane?')] 

recline = recline[complete.cases(recline),] 



recline = separate(data = recline, 
                   col = 'How tall are you?', 
                   into = c("feet", "inches"), sep = "\\'") 

unique(recline$feet) 
# R split the tall column into feet and inches. Feet includes 6 feet, 5 feet, and 'Under 5 feet' 



recline$height = 
  ifelse(recline$feet ==
           "6", "Tall", "Short") 
# Make 6 feet = Tall while anything else (5 feet and Under 5 feet) = Short 



recline$`Is itrude to recline your seat on a plane?` =
  ifelse(recline$`Is itrude to recline your seat on a plane?` ==
           "No, not rude at all", "No", "Yes") 



#Contingency table with being "Tall" with finding reclining rude.   

recline2 = cbind(recline$height, recline$`Is itrude to recline your seat on a plane?`) 

colnames(recline2) = c("Height", "Recline_Rude") 

recline_cont = xtabs(~ Height + Recline_Rude, recline2) 

recline_cont 



Odds_Ratio= (420*65)/(287*82)  

Odds_Ratio #The odds of finding reclining rude is 16% higher if you are tall than if you are short.   



chisq.test(recline_cont, correct=F) # Chi-squared statistic: 0.65953, df = 1, p-value = 0.4167 
#--> not significant 



```

################ Rude to speak to stranger vs reclining ################  
```{r}


recstranger = 
  data[,c('Generally speaking, is it rude to say more than a few words tothe stranger sitting next to you on a plane?', 
                      'Is itrude to recline your seat on a plane?')]  

recstranger= recstranger[complete.cases(recstranger),]  



recstranger$`Generally speaking, is it rude to say more than a few words tothe stranger sitting next to you on a plane?` =
  ifelse(recstranger$`Generally speaking, is it rude to say more than a few words tothe stranger sitting next to you on a plane?` ==
           "No, not at all rude", "No", "Yes")  

recstranger$`Is itrude to recline your seat on a plane?`=
  ifelse(recstranger$`Is itrude to recline your seat on a plane?` ==
           "No, not rude at all", "No", "Yes")  



recst_tab = 
  xtabs(~ recstranger$`Generally speaking, is it rude to say more than a few words tothe stranger sitting next to you on a plane?` +
                    recstranger$`Is itrude to recline your seat on a plane?` )  

recst_tab  



mcnemar.test(recst_tab, correct=F) #test statistic: 86.503, df = 1, p-value < 2.2e-16  

diffpropci.Wald.mp(85, 257, 854, 0.95) #Wald 95% C.I. of difference in marginal probabilities  

##(0.1611693, 0.2416410)  



## Probability of saying it is rude to speak more than a few words to a stranger next to you was 
#16% to 24% higher than saying it is rude to recline your seat.  
```




